\chapter*{Introduction}

In bioinformatics, the evolutionary relationships between entities are expressed with~a~phylogenetic tree depicted by~a~tree from graph theory. We are working with two types of phylogenetic trees in this thesis: species trees and gene trees. A species tree describes the evolutionary events for a set of species and a gene tree represents evolutionary events influencing a single gene within these species. We assume that only evolutionary events of speciation, duplication and gene loss can happen in history. Speciation represents the creation of new distinct species. Duplication occurs when a gene is copied and inserted in another position in the genetic information. The~opposite of~duplication is gene loss, where a gene is lost from genetic information. The~mapping of~the~gene tree to species tree is called the reconciliation, which allows better detection of evolutionary events. Over the years, many reconciliation approaches have been developed. In 2008, an isometric reconciliation was introduced, where branch lengths of both phylogenetic trees are known and considered in mapping. We divide the isometric reconciliation algorithms into two groups: those requiring exact branch lengths and those allowing inexact branch lengths. We discuss the basic terminology and different approaches to solving the problem of reconciliation with concrete solutions to this problem in the form of software in the first chapter.

In this thesis, we aim at studying the problem of isometric reconciliation with~inexact branch lengths since the exact branch lengths are estimated, and thus not exactly known. This field was studied before~\cite{chladek_thesis} with a proposed algorithm to compute a reconciliation with minimizing the duplication and gene loss evolutionary events. The running time of their algorithm is $O(N^4 log N)$. We build on their algorithm and introduce two new algorithms described in more details in Chapter~\ref{Algorithms}.

The proposed algorithms are then implemented into an existing source code from~a~previous thesis~\cite{chladek_thesis}. In Chapter~\ref{Implementation}, we describe the properties of the implemented software with highlighted changes from the previous source code. We define the needed input for our software with an optional setting for the user and subsequent output.

In the last chapter, we experimentally test the implemented software on the simulated and real biological data. We run the software for different settings, evaluate it and compare it with other reconciliation software. We interpret and discuss the results of testing the implemented software.


